Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-11-11 AAAI2026/a_trick.tex
rr

\subsection*{A trick of $L_1$ norm constraints on input}

A trick to set $L_1$ norm constraints on the input space into linear constraints is described below.

Suppose the dimension of input space is $n$, and $x_i,i\leq n$ are the input variables, $\vec{x}$ is the input vector, and we hope to set a constraints of $\vec{x}$ that $\|\vec{x}\| \leq c$,  then we can set the constraints by add variables $A_i,i<n$ with the only constraints \begin{align}\label{L1constraint}\begin{cases}
	A_i \geq x_i &\text{ for }i \leq n\\ 
	A_i\geq -x_i &\text{ for }i \leq n\\
	\Sigma_{i\leq n} A_i \leq c	&\end{cases}
\end{align}

We show that this constraint is equivalent to $\|\vec{x}\| \leq c$.

The first two lines of constraints \ref{L1constraint} is equivalent to that $A_i\geq |x_i|$. So the whole constraints is non-weaker than $\|\vec{x}\| \leq c$.

Next, for every instance that satisfy $\|\vec{x}\| \leq c$, then there is also an instance which satisfy constraints \ref{L1constraint}: simply let all $A_i = |x_i|$. And this shows that $\|\vec{x}\| \leq c$ is non-weaker than constraints \ref{L1constraint}. Therefore they are equivalent.

>>>>

\subsection*{Encoding $L_1$-perturbations with linear variables}

Let $\vec{x}=(x_i)_{i\leq n}$ be an input vector, with $n$ dimensions.

\begin{proposition}
	\label{prop.l1}
Define $n$ linear variables $(A_i)_{i<n}$ with the constraints 
\begin{align}\label{L1constraint}\begin{cases}
	A_i \geq x_i &\text{ for }i \leq n\\ 
	A_i\geq -x_i &\text{ for }i \leq n\\
	\sum_{i\leq n} A_i \leq c	&\end{cases}
\end{align}

This is equivalent with asking $||\vec{x}||_{L_1} = \sum_{i\leq n} |x_i| \leq c$.
\end{proposition}

The proof is simple: the first two lines of constraints (\ref{L1constraint}) are equivalent with $A_i \geq |x_i|$. So the whole constraint (\ref{L1constraint}) 
implies that $\sum_{i\leq n} |x_i| \leq c$.

Conversely, for every instance that satisfies $\sum_{i\leq n} |x_i| \leq c$, then there is also an instance which satisfies constraints (\ref{L1constraint}),  by defining $A_i = |x_i|$ for all $i \leq n$.

This shows that constraints (\ref{L1constraint}) is equivalent with 
$||\vec{x}||_{L_1} = \sum_{i\leq n} |x_i| \leq c$.

### 2025-11-11 AAAI2026/abstract.tex
rr

In this paper, we consider {\em global} robustness, that is, guarantees not restricted to a set of local images. We focus on deriving {\em bounds} on the variation of output values across different decision classes of a DNN under a given perturbation. This global verification problem is significantly more complex than local robustness, as the number of variables doubles (from the deviation image to the image and its deviation). Further, the values each neuron can take are no longer limited to a small neighborhood. To obtain usable bounds, we develop several novel  MILP models for global robustness, with different trade-offs. Last, we use order reduction techniques to reduce the space of inputs considered, avoiding unrealistic images, by using principal component analysis (PCA). This results into the {\em real-time} certification of the robustness for $86\%$ of incoming images in the MNIST benchmark for an $L_1$-perturbation of $.5$, as well as for a surrogate computing the 
%latent
plastic strain associated with a pipe deformation map.

>>>>

In this paper, we consider {\em global} robustness, that is, guarantees not restricted to a set of local images. We focus on deriving {\em bounds} on the variation of output values across different decision classes of a DNN under a given perturbation. This global verification problem is significantly more complex than local robustness, as the number of variables doubles (from the deviation image to the image and its deviation). Further, the values each neuron can take are no longer limited to a small neighborhood. To obtain usable bounds, we develop several novel  MILP models for global robustness, with different trade-offs. Last, we use order reduction techniques to reduce the space of inputs considered, avoiding unrealistic images, by using principal component analysis (PCA). These bounds ensure the {\em real-time} certification of robustness for $86\%$ of incoming images in the MNIST benchmark for an $L_1$-perturbation of $.5$, as well as faitfhulness of a surrogate computing the 
%latent
plastic strain due to the deformation of a pipe.

### 2025-11-11 AAAI2026/related.tex
rr

\section{Related Work}

@@

Closer to our paper, several works  \cite{Marabou,lipshitz,ITNE,GROCET} consider a global certified restricted Lipschitz bound, to understand how the output change with input perturbations. However, they do not consider whether the classification changes.

>>>>

\section{Related Work on Global Robustness}

@@

More related, several works  \cite{Marabou,lipshitz,ITNE,GROCET} consider certifying {\em restricted} Lipschitz bound, to understand how the output can change given an input perturbations. However, they do not consider whether the classification changes.

### 2025-11-11 ATVA 2025/Final_version/introduction.tex
update draft of final version
rr

\begin{table}[t!]
	\vspace{-1.0cm}
	\centering
		\caption{Images verified by $\alpha,\beta$-CROWN with different time-outs (TO) on 7 DNNs, and average runtime per image. The 6 first DNNs are complex instances. The last DNN (ResNet) is an easy instance (trained using Wong to be easy to verify, but with a very low accuracy level), provided for reference.
	}
	\begin{tabular}{||l|c|c|c||c|c|c||}
		\hline
		Network & nbr & Accuracy {$^*$}  & Upper  {$^*$}& $\alpha,\beta$-CROWN& $\alpha,\beta$-CROWN & $\alpha,\beta$-CROWN \\ 
		Perturbation & activ. &  & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
		MNIST 5$\times$100 & 500 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ & ReLU & &  & 6.9s &  18.9s &  1026s  \\  \hline
		MNIST 5$\times$200 & 1000 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & ReLU & &  & 6.5s &  16.6s &  930s  \\  \hline
		MNIST 8$\times$100 & 800 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ & ReLU &  &  & 7.2s &  20.1s &  930s  \\  \hline
		MNIST 8$\times$200 & 1600 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & ReLU & &  & 6.8s &  18.2s &  1083s  \\  \hline
		MNIST 6$\times$500 & 3000 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & ReLU & &  & 6.4s &  16.4s &  1003s  \\  \hline
		CIFAR CNN-B--Adv &  16634  &   78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$& ReLU &  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet &   107496  & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 8/255$ & ReLU &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
	\begin{minipage}{0.9\linewidth}
		\footnotesize
		{\textsuperscript{*} Accuracy means the accuracy of the network itself on test set, and Upper Bound is 100\% the percentage of unsafe examples.} \\
	\end{minipage}
	\label{table_beta}
		\vspace{-0.5cm}
\end{table}

@@

\begin{table}[t!]
	\centering
	\vspace{-0.35cm}
	\caption{Result of non-BaB methods on the hard $5 \times 100$ with TO = 10 000s. 
		%Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and 
		Only NNenum verifies more instances (9\% out of 50\% undecided images) than $\alpha,\beta$-CROWN (40\%), at the cost of a much larger runtime (4995s vs 1026s).
		}
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network &  Accuracy & Upper  & Marabou 2.0 & NNenum &  Full MILP  \\ \hline
		MNIST 5$\times$100 & 99\% & 90\% & 28\% & $49\%$ & 40 \%    \\
		$\epsilon = 0.026$ & &  &6200s &  4995s & 6100s
		  \\  \hline
		%  MNIST 5$\times$100 & 99\% & 90\% & 28\% & $46\%$ & 37 \%    \\
		%  TO=2000s & &  & 1250s  & 1194s & 1466s \\  \hline	
	\end{tabular}
	\vspace{-0.65cm}
\label{table_complete}
\end{table}

>>>>

\begin{table}[t!]
%	\vspace{-1.0cm}
	\centering
	\begin{tabular}{||l|c|c|c||c|c|c||}
		\hline
		Network & nbr & Accur.  & Upper & $\alpha,\beta$-CROWN& $\alpha,\beta$-CROWN & $\alpha,\beta$-CROWN \\ 
		Perturbation & activ. &  & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
		MNIST 5$\times$100 & 500 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ & ReLU & &  & 6.9s &  18.9s &  1026s  \\  \hline
		MNIST 5$\times$200 & 1000 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & ReLU & &  & 6.5s &  16.6s &  930s  \\  \hline
		MNIST 8$\times$100 & 800 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ & ReLU &  &  & 7.2s &  20.1s &  930s  \\  \hline
		MNIST 8$\times$200 & 1600 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & ReLU & &  & 6.8s &  18.2s &  1083s  \\  \hline
		MNIST 6$\times$500 & 3000 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & ReLU & &  & 6.4s &  16.4s &  1003s  \\  \hline
		CIFAR CNN-B--Adv &  16634  &   78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$& ReLU &  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet &   107496  & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 8/255$ & ReLU &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
%	\begin{minipage}{0.9\linewidth}
%		\footnotesize
%		{\textsuperscript{*} Accuracy means the accuracy of the network itself on test set, and Upper Bound is 100\% the percentage of unsafe examples.} \\
%	\end{minipage}
		\caption{Accuracy of DNN (class predicted vs ground truth), upper bound on robustness (robustness attacks found on remaining images), and $\%$ of images verified by $\alpha,\beta$-CROWN with different time-outs (TO) on 7 DNNs, and average runtime per image. The 6 first DNNs are complex instances. The last DNN (ResNet) is an easy instance (trained using Wong to be easy to verify, but with a very low accuracy level), provided for reference.
	}
\label{table_beta}
%		\vspace{-0.5cm}
\end{table}

@@

\begin{table}[b!]
	\centering
%	\vspace{-0.35cm}
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network &  Accuracy & Upper  & Marabou 2.0 & NNenum &  Full MILP  \\ \hline
		MNIST 5$\times$100 & 99\% & 90\% & 28\% & $49\%$ & 40 \%    \\
		$\epsilon = 0.026$ & &  &6200s &  4995s & 6100s
		  \\  \hline
		%  MNIST 5$\times$100 & 99\% & 90\% & 28\% & $46\%$ & 37 \%    \\
		%  TO=2000s & &  & 1250s  & 1194s & 1466s \\  \hline	
	\end{tabular}
%	\vspace{-0.65cm}
	\caption{Result of non-BaB methods on the hard $5 \times 100$ with TO = 10 000s. 
		%Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and 
		Only NNenum verifies more instances (9\% out of 50\% undecided images) than $\alpha,\beta$-CROWN (40\%), at the cost of a much larger runtime (4995s vs 1026s).
		}
\label{table_complete}
\end{table}

### 2025-11-11 lipschitz 2025/Lipschitz_a_proposition.tex
create a new folder
rr

\begin{definition}
	Suppose we have a function $f$ from $\mathbb{R}^n$ to $\mathbb{R}^m$ and $||$ is the $L_\infty$ norm.

	For a number $\epsilon > 0$, $K_\epsilon$ is the smallest number such that for any $d\geq \epsilon$, any inputs $x,y$:\begin{align*}
	 |x-y| \leq d \implies |f(x)-f(y)| \leq dK_\epsilon 
	\end{align*}
\end{definition}

@@

	Continue above definition. Suppose for a number $\epsilon$ we have $K_\epsilon$, then for any number $ \epsilon N\leq d\leq \epsilon(N+1)$ for an integer $N \geq 1$, we have that $$K_d \leq K_\epsilon\frac{N+1}{N}.$$
\end{proposition}

\begin{proof}
By assumption, $d \in [\epsilon N  , \epsilon(N+1)]$ .

	For any $x, y$ that $|x-y| \leq d$, we have $|x-y| \leq \epsilon(N+1) $. Then we can divide the line segment between $x, y$ into $N+1$ pieces: $x_0 = x, x_1, x_2, \cdots, x_{N+1} = y$ such that $|x_i-x_{i+1}| \leq \epsilon$. Then we can apply the definition of $K_\epsilon$ for each pieces.

	Therefore, we have that, for any two inputs $x,y$:
		\begin{align*}
		|x-y| \leq d \implies |f(x)-f(y)| \leq  \epsilon(N+1) K_\epsilon
	\end{align*}

	Hence by the definition, we have that $$K_d <= \dfrac{\epsilon(N+1) K_\epsilon }{d}<=\dfrac{\epsilon(N+1)K_\epsilon}{\epsilon N} = K_\epsilon\frac{N+1}{N}.$$

	This is what we want to show.

\begin{corollary}
	\begin{enumerate}
		\item Continue the definition. Suppose for a number $\epsilon$ we have $K_\epsilon$, then for any number $d\geq \epsilon N$ for an integer $N \geq 1$, we have that $$K_d \leq K_\epsilon\frac{N+1}{N}.$$

		\item Therefore, for any inputs $x,y$ that $|x-y|\geq N\epsilon$ for an integer $N\geq 1$, \begin{align*}
			\dfrac{|f(x)-f(y)|}{|x-y|}\leq K_\epsilon\frac{N+1}{N}
		\end{align*}

		\item 	If for a number $\epsilon$, we get the value of $K_{0.5 \epsilon}$, then we will have a Lipschizt constant $\frac{2+1}{2}K_{0.5 \epsilon}$ working for $|x-y|\geq\epsilon$.

		\item 	Similarly, if for a number $\epsilon$, we get the value of $K_{\epsilon}$, then we will have a Lipschizt constant $\frac{1+1}{1}K_{\epsilon}$ working for $|x-y|\geq\epsilon$.
	\end{enumerate}
\end{corollary}

\begin{proof}
	All three are simply application of the proposition.
\end{proof}

\begin{proposition}
	Continue the definition. Suppose we have a number $K$ that for any $d$ in one of $\epsilon, 1.1\epsilon, 1.2\epsilon, \cdots, 1.9\epsilon$, we have $K_d\leq K$.

	Then, for any two inputs $x,y$,if $|x-y|\geq \epsilon$, then $\dfrac{|f(x)-f(y)|}{|x-y|}\leq 1.1 K$. 
\end{proposition}

\begin{proof}
	Let $x,y$ be any two inputs, we need to show that if $|x-y|\geq \epsilon$, then $\dfrac{|f(x)-f(y)|}{|x-y|}\leq 1.1 K$.

	If $|x-y| \geq 10\epsilon$, this is clear by above corollary part 2. So we need to consider $\epsilon\leq |x-y|< 10\epsilon$.

	For any inputs $x,y$ that $\epsilon\leq|x-y|< 10\epsilon$, there exists a sum $x_1+x_2+\cdots+x_n$ by numbers from (allowing repetitions) $\epsilon, 1.1\epsilon, 1.2\epsilon, \cdots, 1.9\epsilon$ such that \begin{align*}
	 \epsilon \leq x_1+x_2+\cdots+x_n -0.1\epsilon \leq |x-y| \leq x_1+x_2+\cdots+x_n
	\end{align*}

	By assumption, divide the line segment from $x$ to $y$ into pieces according to $x_1, x_2,\cdots,x_n$, then we will have $$|f(x)-f(y)|\leq Kx_1+Kx_2+\cdots+Kx_n.$$

	Hence,\begin{align*}
		\dfrac{|f(x)-f(y)|}{|x-y|} &\leq \dfrac{Kx_1+Kx_2+\cdots+Kx_n}{x_1+x_2+\cdots+x_n -0.1\epsilon}\\
		& = K\cdot( 1+  \dfrac{0.1\epsilon}{x_1+x_2+\cdots+x_n -0.1\epsilon})\\
		& \leq K\cdot( 1+  \dfrac{0.1\epsilon}{\epsilon})\\
		& \leq 1.1 K
	\end{align*}

>>>>

	\begin{definition}[$\varepsilon$-diff bound]
	Suppose we have a function $f$ from $\mathbb{R}^n$ to $\mathbb{R}^m$ and $||$ is the $L_\infty$ norm.

	For a number $\varepsilon\in\mathbb{R}$, an $\varepsilon$-diff bound $D_\varepsilon$ is a number such that for any inputs $x,y$: \begin{align*}
	|x-y|\leq \varepsilon \implies |f(x)-f(y)| \leq D_\varepsilon \cdot \varepsilon
\end{align*}

	\end{definition}

\begin{definition}[Lipschitz above $\varepsilon$ constant ]
	Suppose we have a function $f$ from $\mathbb{R}^n$ to $\mathbb{R}^m$ and $||$ is the $L_\infty$ norm.

	For a number $\varepsilon\in\mathbb{R}$, a Lipschitz above $\varepsilon$ constant  $K_\varepsilon$,  is a number such that for any inputs $x,y$: \begin{align*}
	|x-y|\geq \varepsilon &\implies |f(x)-f(y)| \leq K_\varepsilon \cdot |x-y|\\
	|x-y|<\varepsilon &\implies |f(x)-f(y)| \leq K_\varepsilon \cdot \varepsilon\\
	\end{align*}

\end{definition}

@@

	Suppose $D$ is an $\varepsilon$-diff bound for $f(x)$. Then for any $N\in\mathbb{Z}^+$, $D\frac{N+1}{N}$ is a Lipschitz about $N\varepsilon$ constant.

	In formula, for any two inputs $x,y$, if \begin{align*}
		|x-y|\leq \varepsilon \implies |f(x)-f(y)| \leq D \cdot \varepsilon,
	\end{align*} then 	 \begin{align*}
	|x-y|\geq N\varepsilon &\implies |f(x)-f(y)| \leq D\frac{N+1}{N} \cdot |x-y|\\
	|x-y|<N\varepsilon &\implies |f(x)-f(y)| \leq D\frac{N+1}{N} \cdot N\varepsilon\\
	\end{align*}
\end{proposition}

\begin{proof} We fix the number $N\in\mathbb{Z}^+$ and assume we have two inputs $x, y$.

	The first case, $|x-y|\geq N\varepsilon$. Then we assume $|x-y| \in [M\varepsilon ,  (M+1)\varepsilon]$ for another integer $M\geq N$. Then we can divide the line segment between $x, y$ into $M+1$ pieces: $x_0 = x, x_1, x_2, \cdots, x_{M+1} = y$ such that $|x_i-x_{i+1}| \leq \varepsilon$ and apply the definition of $\varepsilon$-diff bound $D$ for each pieces:\begin{align*}
		|f(x)-f(y)| &= |f(x_{M+1})-f(x_M)+\cdots+f(x_1)-f(x_0)|\\
		&\leq |f(x_{M+1})-f(x_M)|+\cdots+|f(x_1)-f(x_0)|\\
		&\leq D\varepsilon + \cdots +D\varepsilon \\
		&= (M+1)D\varepsilon
	\end{align*}
	Hence\begin{align*}
		|f(x)-f(y)| &\leq (M+1)D\varepsilon \\
		&\leq D\cdot (M+1)\varepsilon \frac{|x-y|}{M\varepsilon}\\
		&= D\cdot\frac{M+1}{M} |x-y|\\
		&\leq   D\cdot\frac{N+1}{N} |x-y|
	\end{align*}
	The second case, $|x-y|< N\varepsilon$. Similarly we can divide the line segment between $x, y$ into $N$ pieces and then $|f(x)-f(y)|\leq D N\varepsilon\leq |f(x)-f(y)|\leq D \frac{N+1}{N} N\varepsilon$.

	This is what we want to show.
\end{proof}

\begin{proposition}
For any $N\in\mathbb{Z}^+$, suppose for any $a$ in $\{\frac{N}{N}\varepsilon,\frac{N+1}{N}\varepsilon,\cdots, \frac{2N-1}{N}\varepsilon\}$, $D$ is an $a$-diff bound for $f(x)$.

Then $D\frac{N+1}{N}$ is a Lipschitz about $\varepsilon$ constant:\begin{align*}
	|x-y|\geq \varepsilon &\implies |f(x)-f(y)| \leq D\frac{N+1}{N} \cdot |x-y|\\
	|x-y|<\varepsilon &\implies |f(x)-f(y)| \leq D\frac{N+1}{N} \cdot \varepsilon\\
\end{align*}
\end{proposition}

\begin{proof}
We fix the number $N\in\mathbb{Z}^+$ and assume we have two inputs $x, y$.

For the case that $|x-y|<\varepsilon$, this is trivial by definition.

For the case that $|x-y|\geq \varepsilon$, there exists a sum $x_1+x_2+\cdots+x_n$ by numbers from (allowing repetitions) $\{\frac{N}{N}\varepsilon,\frac{N+1}{N}\varepsilon,\cdots, \frac{2N-1}{N}\varepsilon\}$ such that \begin{align*}
	 \varepsilon \leq x_1+x_2+\cdots+x_n -\frac{1}{N}\varepsilon \leq |x-y| \leq x_1+x_2+\cdots+x_n
	\end{align*}
	By assumption, divide the line segment from $x$ to $y$ into pieces according to $x_1, x_2,\cdots,x_n$, then we will have $$|f(x)-f(y)|\leq Dx_1+Dx_2+\cdots+Dx_n.$$

	Hence,\begin{align*}
		\dfrac{|f(x)-f(y)|}{|x-y|} &\leq \dfrac{Dx_1+Dx_2+\cdots+Dx_n}{x_1+x_2+\cdots+x_n -\frac{1}{N}\varepsilon}\\
		& = D\cdot( 1+  \dfrac{\frac{1}{N}\varepsilon}{x_1+x_2+\cdots+x_n -\frac{1}{N}\varepsilon})\\
		& \leq D\cdot( 1+  \dfrac{\frac{1}{N}\varepsilon}{\varepsilon})\\
		& = D \frac{N+1}{N} 
	\end{align*}
